\documentclass{standalone}

\begin{document}

\subsection[Image Quality]{Image Quality}\label{SR:quality}

The most powerful image quality evaluator is given by our eyes.
This is true also for SR problems: the final purpose still remains to obtain images that are better visible for human eyes, the so called \emph{visual loss}.
We can however provide some mathematical formulas which allow to quantitative evaluate the image quality.
In both cases we need to establish a relationship between the original image and the produced one.
Thus, we can formulate a quality score only with a reference image.
In SR problems, or more in general in up-sampling problems, we can compare the original HR image with the image obtained by the output of our model.
In this way our quality score would be a measure of similarity between the two images.

The simpler similarity score can be obtained evaluating the peak-signal-to-noise-ratio (PSNR).
This quantity is commonly used to establish the compression lossless of an image and it can be computed as

\begin{equation}
PSNR = 20 \cdot \log_{10}\left( \frac{\max(I)}{\sqrt(MSE)} \right)
\end{equation}
\\
where $\max(I)$ is the maximum value which can be taken by a pixel in the image (in general it should be $1$ or $255$ depending on the image format chosen) and $MSE$ is the Mean Square Error (ref.~\ref{NN:cost}) between the original image and the reconstructed one.
The MSE for an image can be computed as:

\begin{equation}
MSE = \frac{1}{WH} \sum_{i=1}^{W}\sum_{j=1}^{H} \left( I(i, j) - K(i, j) \right)^2
\end{equation}
\\
where $W$, $H$ are width and height of the two images and $I$, $K$ are the original and reconstructed image, respectively.

In other words, the PSNR is the maximum power of the signal over the background noise.
It is expressed in decibel (dB) because the image values ranging in a wide interval and the logarithmic function rearrange the domain.
Thus, we can conclude that high PSNR values are associated to a good reconstruction of the original image.

The PSNR is probably the most common quality score~\cite{psnr_ssim}, but it is not always related to a qualitative visual quality.
Despite it is commonly used as loss function for SR models.

\begin{table}[htbp]
\centering
\begin{tabular}{lccc}
\hline \rowcolor{darkgrayrow}
         & Nearest       & Bicubic        & Lanczos  \\
\hline
PSNR     & 25.118        & 27.254         & 26.566   \\
SSIM     &  0.847        &  0.894         &  0.871   \\
\hline
\end{tabular}
\caption{Image quality scores: PSNR (peak-signal-to-noise-ratio) and SSIM (Structural SIMilarity index).
The values are computed on the image shown in Fig.~\ref{fig:resampling}.
The original image was down-sampled using a Lanczos algorithm and then re-up-sampled using three different algorithms: nearest, bi-cubic and Lanczos interpolations.
For each interpolation algorithm the PSNR and SSIM was evaluated.
As expected the highest scores were obtained using the bi-cubic algorithm, while the worst reconstruction is performed by the nearest algorithm.
}
\label{tab:psnr}
\end{table}

Considering the series of images shown in Fig.~\ref{fig:resampling} we can evaluate the PSNR score starting from a down-sampled image.
Taking the down-sampled image obtained with the Lanczos algorithm we can compare the original image with their up-sampled version given by the three methods (ref. Tab.~\ref{tab:psnr}).
As expected, the lowest PSNR value is achieved by the nearest interpolation method, while the best performances are obtained by the bi-cubic algorithm.
This confirms the wider use of the bi-cubic method in image processing applications.
Moreover, we have to take in account that an increment of $0.25$ in PSNR value corresponds to a visible improvement for human eyes.

A more advanced quality score, commonly used in super resolution image evaluation, is given by the \emph{Structural SIMilarity index} (SSIM).
The SSIM aims to mathematically evaluate the structural similarity between two images, taking into account also the visible improvements seen by human eyes.
The SSIM function can be expressed as

\begin{equation}
SSIM(I, K) = \frac{1}{N}\sum_{i=1}^{N} SSIM(x_{i}, y_{i})
\end{equation}
\\
where $N$ is the number of arbitrary patches which divides the image\footnote{
  Patch dimensions commonly used are $11\times11$ or $8\times8$.
}.
For each patch the SSIM is computed as

\begin{equation}
SSIM(x, y) = \frac{(2\mu_x\mu_y + c_1)(2\sigma_{xy} + c_2)}{ ({\mu_x}^2 + {\mu_y}^2 + c_1)({\sigma_x}^2 + {\sigma_y}^2 + c_2) }
\end{equation}
\\
where $\mu$ and $\sigma$ are the mean and variance of the images, respectively, and $\sigma_{xy}$ represents the covariance.
The $c_1$ and $c_2$ parameters are fixed to avoid mathematical divergences.
Also in this case, higher values of SSIM correspond to high an similarity between the original image and the reconstructed one.

Based on the previous equation, we can highlight a link with the pooling function discussed in~\ref{NN:pooling}.
Also in this case, in fact, we work with a window/kernel moved along the image, which applies a mathematical function on the underlying pixels.
This equivalence suggests an easy implementation of this method with slight modifications of the previous code.

The evaluation of SSIM quality score on the previous up-sampled images (ref. Fig.~\ref{fig:resampling} and Tab.~\ref{tab:psnr}) confirms the results obtained by the PSNR.
Also in this case the worst reconstruction is obtained by the nearest algorithm, while the highest SSIM is obtained by the bi-cubic algorithm.
The gap between SSIM values is smaller than PSNR ones, but this is due to the different domains of the two functions.

\end{document}
